---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Extended two-way fixed effects (ETWFE)

<!-- badges: start -->
<!-- badges: end -->

The goal of **etwfe** is to estimate extended (Mundlak) two-way fixed effects _a
la_ [Wooldridge
(2021)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345). Briefly, Wooldridge proposes a set of saturated interaction effects to overcome
the potential bias problems that arise from using vanilla TWFE in
difference-in-differences designs. The Wooldridge solution is intuitive and
elegant, but is rather tedious and error prone to code up manually. This package
aims to simplify the process by providing convenience functions that do the
work for you. **etwfe** thus provides an R equivalent of the 
[`JWDID`](https://ideas.repec.org/c/boc/bocode/s459114.html) Stata module and, indeed, shares the same core design elements (albeit with some
different internal choices).

_Note:_ While I've tested **ewtfe** against common use cases, the package is
still under early development and should be considered experimental. I plan
(hope) to add some more features and a full test suite at some point, while the
documentation could also be improved. You can help by identifying any
bugs and filing issues.

## Installation

You can install the development version of **etwfe** from [GitHub](https://github.com/):

``` r
# install.packages("remotes")
remotes::install_github("grantmcdermott/etwfe")
```

## Examples

To demonstrate the core functionality of **etwfe**, I'll follow the lead of
`JWDID` in using an example dataset from the **did** package.

```{r mpdta}
# install.packages("did")
data("mpdta", package = "did")
head(mpdta)
```

Now let's see a simple example. Note that the `gref` argument will be
unnecessary in most cases. But we invoke it here explicitly, since the
never-treated group in our dataset take on an unusual value (here: 0).

```{r ex_simple}
library(etwfe)

etwfe(
  fml  = lemp ~ 0,
  gvar = "first.treat", gref = 0,
  tvar = "year",
  data = mpdta,
  vcov = ~countyreal
  )
```

As you can see, the key `etwfe()` function is effectively a wrapper around
`fixest::feols()`. The resulting object is thus fully compatible with other
**fixest** methods and functions like `etable()`. Non-linear models (e.g.
"poisson") are also supported via the family argument.

One of the advantages of ETWFE is that it provides clear theoretical support for
additional control variables. On the downside, these can tricky to code up
because they must be demeaned and then correctly interacted with all of our main
variables of interest. **etwfe** does all of this for you automatically. Here we
add `lpop` as an additional control in our regression.

```{r mod}
mod = 
  etwfe(
    fml  = lemp ~ lpop,
    gvar = "first.treat", gref = 0,
    tvar = "year",
    data = mpdta,
    vcov = ~countyreal
  )
mod
```

The coefficients from an `etwfe()` estimation are not necessarily meaningful in
of themselves. Instead, we probably wish to aggregate them along some dimension of interest (e.g., an event study). A natural way to perform these aggregations
is by calculating marginal effects. The **etwfe** package provides another convenience function for doing this, `emfx()`, which is itself a thin(ish)
wrapper around `marginaleffects::marginaleffects()`

```{r mod_emfx}
# Other type options incl. "simple" (default), "calendar", and "group"
emfx(mod, type = "event")
```

## Acknowledgements

- [Jeffrey Wooldridge](https://twitter.com/jmwooldridge) for the [underlying theory](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345) that drives this package.
- [Laurent Bergé](https://twitter.com/lrberge) ([**fixest**](https://lrberge.github.io/fixest/)) and [Vincent Arel-Bundock](https://twitter.com/VincentAB) ([**marginaleffects**](https://vincentarelbundock.github.io/marginaleffects)) for maintaining the two wonderful R packages that do most of the heavy lifting under the hood here.
- [Fernando Rios-Avila](https://twitter.com/friosavila) for the [`JWDID`](https://ideas.repec.org/c/boc/bocode/s459114.html) Stata module, which has provided a much-appreciated ground truth for checking results and whose elegant design helped inform my own choices for this R equivalent. 