---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  options(width = 100)
)
```

# Extended Two-way Fixed Effects (ETWFE)

<!-- badges: start -->
[![R-CMD-check](https://github.com/grantmcdermott/etwfe/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/grantmcdermott/etwfe/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/etwfe)](https://CRAN.R-project.org/package=etwfe)
<!-- badges: end -->

The goal of **etwfe** is to estimate extended two-way fixed effects _a la_
Wooldridge ([2021](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345),
[2022](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4183726)).  Briefly,
Wooldridge proposes a set of saturated interaction effects to overcome the
potential bias problems of vanilla TWFE in difference-in-differences designs.
The Wooldridge solution is intuitive and elegant, but rather tedious and error
prone to code up manually. The **etwfe** package aims to simplify the process by
providing convenience functions that do the work for you.

Documentation is available on the package [homepage](https://grantmcdermott.com/etwfe/).

## Installation

You can install **etwfe** from CRAN.

```r
install.packages("etwfe")
```

Or, you can grab the development version from GitHub.

``` r
# install.packages("remotes")
remotes::install_github("grantmcdermott/etwfe")
```

## Quickstart example

A detailed walkthrough of **etwfe** is provided in the introductory vignette 
(available [online](https://grantmcdermott.com/etwfe/articles/etwfe.html), or
type `vignette("etwfe")` in your R console). But here's a quickstart example to
demonstrate the basic syntax.

```{r mpdta}
library(etwfe)

# install.packages("did")
data("mpdta", package = "did")
head(mpdta)

# Estimate the model
mod =
  etwfe(
    fml  = lemp ~ lpop, # outcome ~ controls
    tvar = year,        # time variable
    gvar = first.treat, # group variable
    data = mpdta,       # dataset
    vcov = ~countyreal  # vcov adjustment (here: clustered)
    )
mod

# Event-study treatment effects
emfx(mod, type = "event") |>
  summary()
```

## Acknowledgements

- [Jeffrey Wooldridge](https://twitter.com/jmwooldridge) for the [theory](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345) [underlying](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4183726) ETWFE.
- [Laurent Bergé](https://twitter.com/lrberge) ([**fixest**](https://lrberge.github.io/fixest/)) and [Vincent Arel-Bundock](https://twitter.com/VincentAB) ([**marginaleffects**](https://vincentarelbundock.github.io/marginaleffects/)) for maintaining the two wonderful R packages that do most of the heavy lifting under the hood here.
- [Fernando Rios-Avila](https://twitter.com/friosavila) for the [`JWDID`](https://ideas.repec.org/c/boc/bocode/s459114.html) Stata module, which has provided a welcome foil for unit testing and whose elegant design helped inform my own choices for this R equivalent. 
