---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Extended two-way fixed effects (ETWFE)

<!-- badges: start -->
[![R-CMD-check](https://github.com/grantmcdermott/etwfe/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/grantmcdermott/etwfe/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of **etwfe** is to estimate extended (Mundlak) two-way fixed effects _a
la_ [Wooldridge (2021)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345). 
Briefly, Wooldridge proposes a set of saturated interaction effects to overcome
the potential bias problems that arise from using vanilla TWFE in
difference-in-differences designs. The Wooldridge solution is intuitive and
elegant, but is rather tedious and error prone to code up manually. This package
aims to simplify the process by providing convenience functions that do the
work for you. **etwfe** thus provides an R equivalent of the 
[`JWDID`](https://ideas.repec.org/c/boc/bocode/s459114.html) Stata module and, 
indeed, shares some of the core design elements (albeit with some internal 
differences).

_Note:_ While I've tested **ewtfe** against common use cases, the package is
still under early development and should be considered experimental. I plan
(hope) to add some more features and the documentation could also be improved.
You can help by identifying any bugs and filing issues.

## Installation

You can install the development version of **etwfe** from 
[GitHub](https://github.com/):

``` r
# install.packages("remotes")
remotes::install_github("grantmcdermott/etwfe")
```

## Examples

To demonstrate the core functionality of **etwfe**, I'll follow the lead of
`JWDID` by using an example dataset from the **did** package.

```{r mpdta}
# install.packages("did")
data("mpdta", package = "did")
head(mpdta)
```

Now let's see a simple example. Note that the `gref` argument will be
unnecessary in most cases. But we invoke it here explicitly, since the
"never-treated" group in the `mpdta` dataset takes on an unusual value (here:
0).

```{r mod}
library(etwfe)

mod = 
  etwfe(
    fml  = lemp ~ lpop,           # outcome ~ controls
    tvar = year,                  # time variable
    gvar = first.treat, gref = 0, # group variable (with bespoke ref. level)
    data = mpdta,                 # dataset
    vcov = ~countyreal            # vcov adjustment (here: clustered)
    )
mod
```

As you can see, the key `etwfe()` function is effectively a wrapper around
`fixest::feols()`. (Although, non-linear models are also supported via the
`family` argument.) The resulting object is thus fully compatible with other
**fixest** methods and functions like `etable()`.

```{r etab}
fixest::etable(mod, signif.code = NA)
```

While everyone likes a nice regression table, the raw coefficients from an
`etwfe()` estimation are not necessarily meaningful in of themselves. Instead,
we probably want to aggregate them along some dimension of interest (e.g., an
event study). A natural way to perform these aggregations is by calculating
marginal effects. The **etwfe** package provides another convenience function
for doing this, `emfx()`, which is itself a thin(ish) wrapper around
`marginaleffects::marginaleffects()`

```{r mod_emfx}
# Other type options incl. "simple" (default), "calendar", and "group"
emfx(mod, type = "event")
```

## Acknowledgements

- [Jeffrey Wooldridge](https://twitter.com/jmwooldridge) for the [underlying theory](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345) that drives this package.
- [Laurent Bergé](https://twitter.com/lrberge) ([**fixest**](https://lrberge.github.io/fixest/)) and [Vincent Arel-Bundock](https://twitter.com/VincentAB) ([**marginaleffects**](https://vincentarelbundock.github.io/marginaleffects)) for maintaining the two wonderful R packages that do most of the heavy lifting under the hood here.
- [Fernando Rios-Avila](https://twitter.com/friosavila) for the [`JWDID`](https://ideas.repec.org/c/boc/bocode/s459114.html) Stata module, which has provided a welcome foil for checking results and whose elegant design helped inform my own choices for this R equivalent. 
