---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  options(width = 100)
)
```

# Extended two-way fixed effects (ETWFE)

<!-- badges: start -->
[![R-CMD-check](https://github.com/grantmcdermott/etwfe/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/grantmcdermott/etwfe/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of **etwfe** is to estimate extended (Mundlak) two-way fixed effects _a
la_ [Wooldridge (2021)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345). 
Briefly, Wooldridge proposes a set of saturated interaction effects to overcome
the potential bias problems of vanilla TWFE in
difference-in-differences designs. The Wooldridge solution is intuitive and
elegant, but rather tedious and error prone to code up manually. This package
aims to simplify the process by providing convenience functions that do the
work for you. **etwfe** thus provides an R equivalent of the 
[`JWDID`](https://ideas.repec.org/c/boc/bocode/s459114.html) Stata module and, 
indeed, shares some of the core design elements (albeit with some internal 
differences).

While I've tested **ewtfe** against common use cases, please note that the package is
still under early development and should be considered experimental. I plan
(hope) to add some more features and the documentation could also be improved.
You can help by identifying any bugs and filing issues.

## Installation

You can install the development version of **etwfe** from 
[GitHub](https://github.com/):

``` r
# install.packages("remotes")
remotes::install_github("grantmcdermott/etwfe")
```

_Note:_ **etwfe** relies on the current development versions of **fixest** and
**marginaleffects**. So you'll have to compile **fixest**'s C++ source code on
your system during installation.[^1] Once these dependencies make their way 
to CRAN, I'll submit **etwfe** to CRAN as well so that binaries are available 
for easy install.

## Examples

To demonstrate the core functionality of **etwfe**, we'll use the `mpdta` 
dataset from the **did** package. (You'll need to install the latter 
separately.)

```{r mpdta}
# install.packages("did")
data("mpdta", package = "did")
head(mpdta)
```

Now let's see a simple example.[^2]

```{r mod}
library(etwfe)

mod = 
  etwfe(
    fml  = lemp ~ lpop, # outcome ~ controls
    tvar = year,        # time variable
    gvar = first.treat, # group variable (with bespoke ref. level)
    data = mpdta,       # dataset
    vcov = ~countyreal  # vcov adjustment (here: clustered)
    )
mod
```

As you can see, the key `etwfe()` function is effectively a wrapper around
`fixest::feols()`. (Nonlinear models are also supported via the
`family` argument.) The resulting object is thus fully compatible with other
**fixest** methods and functions like `etable()`.

```{r etab}
fixest::etable(mod, signif.code = NA)
```

While everyone likes a nice regression table, the raw coefficients from an
`etwfe()` estimation are not necessarily meaningful in of themselves. Instead,
we probably want to aggregate them along some dimension of interest (e.g., an
event study). A natural way to perform these aggregations is by calculating
marginal effects. The **etwfe** package provides another convenience function
for doing this, `emfx()`, which is itself a thin(ish) wrapper around
`marginaleffects::marginaleffects()`.

```{r mod_emfx}
# Other type options incl. "simple" (default), "calendar", and "group"
emfx(mod, type = "event")
```

## Acknowledgements

- [Jeffrey Wooldridge](https://twitter.com/jmwooldridge) for the [underlying theory](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345) behind ETWFE.
- [Laurent Bergé](https://twitter.com/lrberge) ([**fixest**](https://lrberge.github.io/fixest/)) and [Vincent Arel-Bundock](https://twitter.com/VincentAB) ([**marginaleffects**](https://vincentarelbundock.github.io/marginaleffects)) for maintaining the two wonderful R packages that do most of the heavy lifting under the hood here.
- [Fernando Rios-Avila](https://twitter.com/friosavila) for the [`JWDID`](https://ideas.repec.org/c/boc/bocode/s459114.html) Stata module, which has provided a welcome foil for unit testing and whose elegant design helped inform my own choices for this R equivalent. 

[^1]: This means you need
[Rtools](https://cran.r-project.org/bin/windows/Rtools) if you're on Windows, 
and [Xcode](https://mac.r-project.org/tools/) if you're on a Mac. Linux users 
should be good to go without any other requirements.

[^2]: See the `?etwfe` helpfile for information about other function arguments 
that can be used to customize the underlying estimation.