---
title: "Introduction to etwfe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to etwfe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
img {
    max-width: 100%;
    max-height: 100%;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)

fixest::setFixest_notes(FALSE)
```

## Background

A canonical research design for social scientists is the so-called
"difference-in-differences" (DiD) design. The essential idea behind DiD is that
we can estimate the causal impact of an intervention ("treatment") by comparing
the pre- and post-intervention outcomes of indivduals that received treatment
("treated" group) against the outcomes of comparable indivduals that didn't
receive treatment ("control" group).^[Good textbook introductions to DiD
are available [here](https://theeffectbook.net/ch-DifferenceinDifference.html)
and [here](https://mixtape.scunning.com/09-difference_in_differences), among
many other places.]

In the classic 2x2 DiD case (two units, two periods), a
simple interaction effect between two dummy variables suffices to recover the 
treatment effect. In base R this might look something like:

```r
lm(y ~ Dtreated_unit * Dpost_treatment, data = somedata)
```

where the resulting coefficient on the `Dtreated_unitTRUE:Dpost_treatmentTRUE`
interaction term represents the treatment effect.

Rather than manually specify the interaction term, in practice researchers
often use an equivalent formulation known as _two-way fixed effects_ (TWFE). The
core idea of TWFE is that we can subsume the interaction term from the previous
code chunk by adding unit and time fixed effects. A single treatment dummy can
then be used to capture the effect of treatment directly. A TWFE regression in
base R might look as follows:

```r
lm(y ~ Dtreat + factor(id) + factor(period), data = somedata)
```

where the treatment effect is now captured by the coefficient on the `Dtreat` 
dummy.

The TWFE shortcut is especially nice for more complicated panel data settings 
with multiple units and multiple times periods. Speaking of which, if you prefer
to use a dedicated fixed effects / panel data package like **fixest**, you could
also estimate the previous regression like so:

```r
library(fixest)
feols(y ~ Dtreat | id + period, data = somedata)
```

These TWFE regressions are easy to run and intuitive, and for a long time
everyone was happy. But it was too good to last. A cottage industry of clever
research now demonstrates that things are not quite so simple. Among other
things, the standard TWFE formulation can impose strange (negative) weighting 
conditions on key parts of the estimation procedure. One implication is that you
risk a high probability of estimate bias in the presence of staggered treatment
rollouts, which are very common in real-life applications.

Fortunately, just as econometricians were taking away one of our
favourite tools, they were kind enough to replace it with some new ones. 
Among these, the proposed approach by 
Wooldridge ([2021](https://dx.doi.org/10.2139/ssrn.3906345),
[2022](https://dx.doi.org/10.2139/ssrn.4183726))
is noteworthy. His idea might be paraphrased as stating that the problem 
with TWFE is not that we were doing it in the first place. Rather, it's that we
weren't doing it enough. Instead of only including a single treatment × time
interaction, Wooldridge recommends that we saturate our model with all possible
interactions between treatment and time variables, including treatment cohorts,
as well as other covariates. He goes on to show that this approach actually
draws an equivalence between different types of estimators (pooled OLS, twoway
Mundlak regression, etc.) So it's not entirely clear what to call it.  But
Wooldridge refers to the general idea as as _extended_ TWFE---or, ETWFE---which
I rather like and is where this package takes its name.

The Wooldridge ETWFE solution is intuitive and elegant. But it is also rather
tedious and error prone to code up manually. You have to correctly specify all
the possible interactions, demean control variables within groups, and then 
recover the treatment effects of interest via an appropriate marginal effect
aggregation. The **etwfe** package aims to simplify the process by providing 
convenience functions that do all this work for you.

## Dataset

To demonstrate the core functionality of **etwfe**, we’ll use the
[`mpdta`](https://bcallaway11.github.io/did/reference/mpdta.html) dataset on 
US teen employment from the **did** package (which you’ll need to install
separately).

```{r}
# install.packages("did")
data("mpdta", package = "did")
head(mpdta)
```

"Treatment" in this dataset refers to an increase in the minimum wage rate. In
the examples that follow, our goal is to estimate the effect of this minimum
wage treatment (`treat`) on the log of teen employment (`lemp`). Notice that the
panel ID is at the county level (`countyreal`), but treatment was staggered
across cohorts (`first.treat`) so that a group of counties were treated at the
same time. In addition to these staggered treatment effects, we also observe log
population (`lpop`) as a potential control variable.

## Basic usage

Let's load **etwfe** and work through its basic functionality. As we'll see,
the core workflow of the package involves two consecutive function calls: 1)
`etwfe()` and 2) `emfx()`.

### `etwfe`

Given the package name, it won't surprise you to learn that the key estimating
function is `etwfe()`. Here's how it would look for our example dataset.

```{r}
library(etwfe)

mod =
  etwfe(
    fml  = lemp ~ lpop, # outcome ~ controls
    tvar = year,        # time variable
    gvar = first.treat, # group variable
    data = mpdta,       # dataset
    vcov = ~countyreal  # vcov adjustment (here: clustered)
    )
```

There are a few things to say about our `etwfe()` argument choices and other
function options, but we'll leave those details aside until a bit later. Right
now, just know that all of the above arguments are required except `vcov`
(though I generally recommend it too, since we probably want to cluster our
standard errors at the individual unit level). 

Let's take a look at our model object.
```{r}
mod
```

What `etwfe()` has done underneath the hood is construct a treatment dummy
variable `.Dtreat` and saturated it together with the other variables of
interest as a set of multiway interaction terms.^[It has also demeaned the 
`lpop` control variable, but that again is something we'll
get back too later. It's not particularly important for interpreting the final
results.] 

You may have noticed that our `etwfe()` call returns a standard 
[**fixest**](https://lrberge.github.io/fixest/) object, since this is what it
uses to perform the underlying estimation. All of the associated methods and
functions from the
**fixest** package are thus compatible with our model object. For example, we
could plot the raw regression coefficients with 
`fixest::coefplot()`, or print them to a nice regression table with
`fixest::etable()`.  However, the raw coefficients from an `etwfe()` estimation
are not particularly meaningful in of themselves. Recall that these are complex,
multiway interaction terms that are probably hard to to interpret on their own.
This insight leads us to our next key function...

### `emfx`

If the raw `etwfe` coefficients aren't particularly useful by themselves, what 
can we do with them instead? Well, we probably want to aggregate them along some 
dimension of interest
(e.g., by groups or time, or as an event study). A natural way to perform these
aggregations is by recovering the appropriate marginal effects. The **etwfe**
package provides another convenience function for doing so: `emfx()`, which is a
thin(ish) wrapper around
[`marginaleffects::slopes()`](https://marginaleffects.com/vignettes/slopes.html).

For example, we can recover the average treatment effect on the treated (ATT) as
follows.

```{r}
emfx(mod)
```

In other words, our model is telling us that an increase in the minimum wage leads 
to an approximate 5 percent decrease in teen employment. 

Beyond simple ATTs, `emfx()` also supports other types of aggregations via the
`type` argument. For example, we can use `type = "calendar"` to get ATTs by
period, or `type = "group"` to get ATTs by cohort groups. But the option that
will probably be useful to most people is `type = "event"`, which will recover
dynamic treatment effects _a la_ an event study. Let's try this out and then
save the resulting object, since I plan to reuse it in a moment.

```{r}
mod_es = emfx(mod, type = "event")
mod_es
```

Our event study suggests that the teen disemployment effect of a minimum wage 
hike is fairly modest at first (3%), but increases over the next few years 
(>10%). In the next section, we'll look at ways to communicate this kind of 
finding to your audience.

### Presentation

Since `emfx()` produces a standard `marginaleffects` object, we can pass it on to
other supported methods and packages. For example, we can pass it on to
[**modelsummary**](https://modelsummary.com/index.html)
to get a nice regression table of the event study coefficients. Note the use of
the `shape` and `coef_rename` arguments below; these are optional but help to
make the output look a bit nicer.

```{r}
library(modelsummary)

# Quick renaming function to replace ".Dtreat" with something more meaningful
rename_fn = function(old_names) {
  new_names = gsub(".Dtreat", "Years post treatment =", old_names)
  setNames(new_names, old_names)
}

modelsummary(
  mod_es,
  shape       = term:event:statistic ~ model,
  coef_rename = rename_fn,
  gof_omit    = "Adj|Within|IC|RMSE",
  title       = "Event study",
  notes       = "Std. errors are clustered at the county level"
)
```

For visualization, you can pass it on to your preferred plotting method. For
example:

```{r}
library(ggplot2)
theme_set(
  theme_minimal() + theme(panel.grid.minor = element_blank())
)

ggplot(mod_es, aes(x = event, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(col = "darkcyan") +
  labs(x = "Years post treatment", y = "Effect on log teen employment")
```

Note that `emfx` only reports post-treatment effects. All pre-treatment effects
are swept out of the estimation because of the way that ETWFE is set up. In
fact, all pre-treatment effects are mechanistically set to zero. This means that
ETWFE cannot be used for interrogating pre-treatment fit (say, a visual
inspection for parallel pre-trends). Still, you can get these zero
pre-treatment effects by changing the `post_only` argument.  I emphasize that
doing so is strictly performative---again, pre-treatment effects are zero by
estimation design---but it might make your event study plot more aesthetically
pleasing.

```{r}
# Use post_only = FALSE to get the "zero" pre-treatment effects
mod_es2 = emfx(mod, type = "event", post_only = FALSE)

ggplot(mod_es2, aes(x = event, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = -1, lty = 2) +
  geom_pointrange(col = "darkcyan") +
  labs(
    x = "Years post treatment", y = "Effect on log teen employment",
    caption = "Note: Zero pre-treatment effects for illustrative purposes only."
  )
```

## Heterogeneous treatment effects

So far we've limited ourselves to homogeneous treatment effects, where the
impact of treatment (i.e., minimum wage hike) is averaged across all US counties
in our dataset. However, many research problems require us to estimate treatment
effects separately across groups and then, potentially, test for differences
between them. For example, we might want to test whether the efficacy of a new
vaccine differs across age groups, or whether a marketing campaign was equally
successful across genders. The ETWFE framework naturally lends itself to these
kinds of heterogeneous treatment effects.

Consider the following example, where we first create a logical dummy variable
for all US counties in the eight
[Great Lake states (GLS)](https://en.wikipedia.org/wiki/Great_Lakes_region).

```{r}
gls_fips = c("IL" = 17, "IN" = 18, "MI" = 26, "MN" = 27,
             "NY" = 36, "OH" = 39, "PA" = 42, "WI" = 55)

mpdta$gls = substr(mpdta$countyreal, 1, 2) %in% gls_fips
```

Now imagine that we are interested in estimating separate treatment effects
for GLS versus non-GLS counties. We do this simply by invoking the optional
`xvar` argument as part of our `etwfe()` call.^[Note that the "x" prefix in 
"xvar" represents a covariate that is *interacted* (×) with treatment, as
opposed to a regular control variable (which could obviously be included as part
of the `fml` RHS.] Any subsequent `emfx()` call on this object will
automatically recognize that we want to recover treatment effects by these two
distinct groups.

```{r}
hmod = etwfe(
   lemp ~ lpop, tvar = year, gvar = first.treat, data = mpdta, 
   vcov = ~countyreal,
   xvar = gls           ## <= het. TEs by gls
   )

# Heterogeneous ATTs (could also specify `type = "event"`, etc.) 
emfx(hmod)
```

The above point estimates might tempt us to think that minimum wage hikes caused
less teen disemployment in GLS counties than in the rest of the US on average.
However, to test this formally we can invoke the powerful
[`hypothesis`](https://marginaleffects.com/vignettes/hypothesis.html)
infrastructure of the underlying **marginaleffects** package. Probably the
easiest way to do this is by using `b[i]`-style positional arguments, where
"`[i]`" denotes the row of the `emfx()` return object. Thus, by specifying
`hypothesis = "b1 = b2"`, we can test whether the ATTs from row 1 (non-GLS) and
row 2 (GLS) are different from one another.

```{r}
emfx(hmod, hypothesis = "b1 = b2")
```

Here we see that there is actually no statistical difference in the average
disemployment effect between GLS and non-GLS counties.

One final aside is that you can easily display the results of heterogeneous 
treatment effects in plot or table form. Here's an example of the latter, where
we again make use of the `modelsummary(..., shape = ...)` argument.

```{r}
modelsummary(
    models      = list("GLS county" = emfx(hmod)),
    shape       = term + statistic ~ model + gls, # add xvar variable (here: gls)
    coef_map    = c(".Dtreat" = "ATT"),
    gof_map     = NA,
    title       = "Comparing the ATT on GLS and non-GLS counties"
)
```

While the simple example here has been limited to a binary comparison of group  
ATTs, note the same logic carries over to richer settings. We can use
the exact same workflow to estimate heterogeneous treatment effects by different
aggregations (e.g., event studies) and across groups with many levels. 

## Other families

Another key feature of the ETWFE approach---one that sets it apart from other
advanced DiD implementations and extensions---is that it supports nonlinear
model (distribution / link) families. Users need simply invoke the `family`
argument. Here's a brief example, where we recast our earlier event-study as a
Poisson regression.

```{r, warning=FALSE, message=FALSE}
mpdta$emp = exp(mpdta$lemp)

etwfe(
  emp ~ lpop, tvar = year, gvar = first.treat, data = mpdta, vcov = ~countyreal,
  family = "poisson"
  ) |>
  emfx("event")
```

## Performance tips

Thinking of the **etwfe** workflow again as a pair of consecutive functional
calls, the first `etwfe()` stage tends to be very fast. We're leveraging the
incredible performance of **fixest** and also taking some shortcuts to avoid
wasting time on nuisance parameters. See the
[Regarding fixed effects](#regarding-fixed-effects) section below for more 
details about this.

For its part, the second `emfx()` stage also tends to be pretty performant. If
your data has less than 100k rows, it's unlikely that you'll have to wait more
than a few seconds to obtain results. However, `emfx`'s computation time does
tend to scale non-linearly with the size of the original data, as well as the
number of interactions from the underlying `etwfe` model object. Without getting
too deep into the weeds, we are relying on a numerical delta method of the
(excellent) **marginaleffects** package underneath the hood to recover the ATTs
of interest. This method requires estimating two prediction models for *each*
coefficient in the model and then computing their standard errors. So it's a
potentially expensive operation that can push the computation time for large
datasets (> 1m rows) up to several minutes or longer.

Fortunately, there are two complementary strategies that you can use to speed
things up. The first is to turn off the most expensive part of the whole
procedure---standard error calculation---by calling `emfx(..., vcov = FALSE)`.
Doing so should bring the estimation time back down to a few seconds or less,
even for datasets in excess of a million rows. Of course, the loss of standard
errors might not be an acceptable trade-off for projects where statistical
inference is critical. But the good news is this first strategy can still be
combined our second strategy: it turns out that collapsing the data by groups
prior to estimating the marginal effects can yield substantial speed gains on
its own. Users can do this by invoking the `emfx(..., collapse = TRUE)`
argument. While the effect here is not as dramatic as the first strategy,
collapsing the data does have the virtue of retaining information about the
standard errors. The trade-off this time, however, is that collapsing our data
does lead to a loss in accuracy for our estimated parameters. On the other hand,
testing suggests that this loss in accuracy tends to be relatively minor, with
results equivalent up to the 1st or 2nd significant decimal place (or even
better).

Summarizing, here is a quick plan of attack for you to try if you are worried
about the estimation time for large datasets and models:

0. Estimate `mod = etwfe(...)` as per usual.
1. Run `emfx(mod, vcov = FALSE, ...)`.
2. Run `emfx(mod, vcov = FALSE, collapse = TRUE, ...)`.
3. Compare the point estimates from steps 1 and 2. If they are are similar 
enough to your satisfaction, get the approximate standard errors by running 
`emfx(mod, collapse = TRUE, ...)`.

It's a bit of performance art, since all of the examples in this vignette
complete very quickly anyway. But here is a reworking of our earlier event study
example to demonstrate this performance-conscious workflow.

```{r}
# Step 0 already complete: using the same `mod` object from earlier...

# Step 1
emfx(mod, type = "event", vcov = FALSE)

# Step 2
emfx(mod, type = "event", vcov = FALSE, collapse = TRUE)

# Step 3: Results from 1 and 2 are similar enough, so get approx. SEs
mod_es2 = emfx(mod, type = "event", collapse = TRUE)
```

To put a fine point on it, we can can compare our original event study with the
collapsed estimates and see that the results are indeed very similar.

```{r}
modelsummary(
    list("Original" = mod_es, "Collapsed" = mod_es2),
    shape       = term:event:statistic ~ model,
    coef_rename = rename_fn,
    gof_omit    = "Adj|Within|IC|RMSE",
    title       = "Event study",
    notes       = "Std. errors are clustered at the county level"
)
```

## Under the hood

Now that you've seen **etwfe** in action, let's circle back to what the package
is doing under the hood. This section isn't necessary for you to use the
package; feel free to skip it. But a review of the internal details should help
you to optimize for different scenarios and also give you a better understanding
of **etwfe's** default choices. 

### Manual implementation

As I keep reiterating, the ETWFE approach basically involves saturating the
regression with interaction effects. You can easily grab the formula of an
estimated model to see for yourself.

```{r}
mod$fml_all
```

At this point, however, you may notice a few things. The first is that our
formula references several variables that aren't in the original dataset. An
obvious one is the `.Dtreat` treatment dummy. A more subtle one is `lpop_dm`,
which is the _demeaned_ (i.e., group-centered) version of our `lpop` control
variable.  All control variables have to be demeaned before
they are interacted in the ETWFE setting. Here's how you could have constructed
the dataset ahead of time and estimated the ETWFE regression manually:

```{r}
# First construct the dataset
mpdta2 = mpdta |>
  transform(
    .Dtreat = year >= first.treat & first.treat != 0,
    lpop_dm = ave(lpop, first.treat, year, FUN = \(x) x - mean(x, na.rm = TRUE))
  )

# Then estimate the manual version of etwfe
mod2 = fixest::feols(
  lemp ~ .Dtreat:i(first.treat, i.year, ref = 0, ref2 = 2003) / lpop_dm |
    first.treat[lpop] + year[lpop],
  data = mpdta2,
  vcov = ~countyreal
)
```

We can confirm that the manual approach yields the same output as our original
etwfe regression. I'll use **modelsummary** to do that here, since I've already
loaded it above.^[Another option would be to use `fixest::etable()`].

```{r}
modelsummary(
  list("etwfe" = mod, "manual" = mod2),
  gof_map = NA # drop all goodness-of-fit info for brevity
)
```

To transform these raw coefficients into their more meaningful ATT counterparts,
we just need to perform the appropriate marginal effects operation. For example,
here's how we can get both the simple ATTs and event-study ATTs from earlier. 
This is what `emfx()` is doing behind the scenes.

```{r}
library(marginaleffects)

# Simple ATT
slopes(
  mod2, 
  newdata   = subset(mpdta2, .Dtreat), # we only want rows where .Dtreat is TRUE
  variables = ".Dtreat", 
  by        = ".Dtreat"
  )

# Event study
slopes(
  mod2, 
  newdata   = transform(subset(mpdta2, .Dtreat), event = year - first.treat),
  variables = ".Dtreat", 
  by        = "event"
  )
```

### Regarding fixed effects

Let's switch gears and talk about fixed effects quickly. If you are a regular 
**fixest** user, you may have noticed that we've been invoking its 
[varying
slopes](https://lrberge.github.io/fixest/articles/fixest_walkthrough.html#varying-slopes-fex)
syntax in the fixed effect slot (i.e., `first.treat[lpop]` and `year[lpop]`).
The reason for this is part practical, part philosophical. From a practical
perspective, `factor_var[numeric_var]` is equivalent to base R's 
`factor_var/numeric_var` "nesting" syntax but is much faster for high-dimensional 
factors.^[We won't see a speed-up for this small dataset, but it can make a
significant difference for large datasets.] From a philosophical perspective,
**etwfe** tries to limit the amount of extraneous information that it reports
to users. Most of the interaction effects in the ETWFE framework are just acting
as controls. By relegating them to the fixed effects slot, we can avoid polluting
the user's console with a host of extra coefficients. Nonetheless, we can control 
this behaviour with the `fe` ("fixed effects") argument. Consider the following
options and their manual equivalents.

```{r, message=FALSE, warning=FALSE}
# fe = "feo" (fixed effects only)
mod_feo = etwfe(
  lemp ~ lpop, tvar = year, gvar = first.treat, data = mpdta, vcov = ~countyreal,
  fe = "feo"
)
# ... which is equivalent to the manual regression
mod_feo2 = fixest::feols(
  lemp ~ .Dtreat:i(first.treat, i.year, ref = 0, ref2 = 2003) / lpop_dm +
    lpop + i(first.treat, lpop, ref = 0) + i(year, lpop, ref = 2003) |
    first.treat + year,
  data = mpdta2, vcov = ~countyreal
)

# fe = "none"
mod_none = etwfe(
  lemp ~ lpop, tvar = year, gvar = first.treat, data = mpdta, vcov = ~countyreal,
  fe = "none"
)
# ... which is equivalent to the manual regression
mod_none2 = fixest::feols(
  lemp ~ .Dtreat:i(first.treat, i.year, ref = 0, ref2 = 2003) / lpop_dm +
    lpop + i(first.treat, lpop, ref = 0) + i(year, lpop, ref = 2003) +
    i(first.treat, ref = 0) + i(year, ref = 2003),
  data = mpdta2, vcov = ~countyreal
)
```

I'll leave it up to you to pass any of these models to `emfx` to confirm that
they give correct aggregated treatment effects. But we can quickly demonstrate
in a regression table that they all return the same raw coefficients.

```{r}
mods = list(
  "etwfe"         = mod,
  "manual"        = mod2,
  "etwfe (feo)"   = mod_feo,
  "manual (feo)"  = mod_feo2,
  "etwfe (none)"  = mod_none,
  "manual (none)" = mod_none2
)

modelsummary(mods, gof_map = NA)
```

A final point to note about fixed effects is that **etwfe** defaults to using
group-level (i.e., cohort-level) fixed effects like `first.treat`, rather than
unit-level fixed effects like `countyreal`. This design decision reflects a neat
ancillary result in Wooldridge (2021), which proves the equivalence between the
two types of fixed effects for linear cases. Group-level effects have the virtue
of being faster to estimate, since there are fewer factor levels. Moreover, they
are _required_ for nonlinear model families like Poisson per the underlying
ETWFE theory. Still, you can specify unit-level fixed effects for the linear
case through the `ivar` argument. Again, we can easily confirm that this yields
the same estimated treatment effects as the group-level default (although the
standard errors will be slightly different).

```{r}
mod_es_i = etwfe(
  lemp ~ lpop, tvar = year, gvar = first.treat, data = mpdta,
  ivar = countyreal  # NEW: Use unit-level (county) FEs
  ) |>
  emfx("event")

modelsummary(
  list("Group-level FEs (default)" = mod_es, "Unit-level FEs" = mod_es_i),
  shape       = term:event:statistic ~ model,
  coef_rename = rename_fn,
  gof_omit    = "Adj|Within|IC|RMSE"
)
```
