---
title: "Introduction to etwfe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{etwfe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
fixest::setFixest_notes(FALSE)
```

## Background

A canonical research design for social scientists is the so-called
"differences-in-differences" (DiD) design.^[Good textbook introductions to DiD
are available [here](https://theeffectbook.net/ch-DifferenceinDifference.html)
and [here](https://mixtape.scunning.com/09-difference_in_differences), among
many other places.] In the classic 2x2 DiD case (two units, two periods), a
simple interaction effect between two dummy variables suffices to recover the 
treatment effect. In base R this might look something like:

```r
lm(y ~ Dtreated_unit * Dpost_treatment, data = somedata)
```

In practice, rather than specify the interaction directly, researchers often use
an equivalent formulation known as _two-way fixed effects_ (TWFE). The core idea
of TWFE is that we can subsume our dummy variables above (`Dtreated_unit` and
`Dpost_treatment`) by adding unit and time fixed effects to the regression. A
single treatment dummy can then be used to capture the effect of treatment
directly. This shortcut is especially nice for more complicated panel data
settings with multiple units and multiple times periods. A TWFE in base R might
look as follows:

```r
lm(y ~ Dtreat + factor(id) + factor(period), data = somedata)
```

This type of regression is easy to run and intuitive, and for a long time
everyone was happy. But it was too good to last. A cottage industry of smart
research now demonstrates that things aren't quite so simple. Among other
things, the core TWFE result is very likely to be biased if the rollout of
treatment was staggered over time (a common occurrence).

Fortunately, just as econometricians were taking away one of our
favourite tools, they were kind enough to replace it with some new ones. 
Among these, the proposed approach by 
Wooldridge ([2021](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3906345),
[2022](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4183726))
is noteworthy. His idea might be paraphrased as stating that the problem 
with TWFE is not that we were doing it in the first place. Rather, it's that we
weren't doing it enough. Instead of only including a single treatment × time
interaction, Wooldridge recommends that we saturate our model with all possible
interactions between treatment and time variables, including treatment cohorts,
as well as other covariates. He goes on to show that this approach actually
draws an equivalence between different types of estimators (pooled OLS, twoway
Mundlak regression, etc.) So it's not entirely clear what to call it.  But
Wooldridge refers to the general idea as as _extended_ TWFE---or, ETWFE---which
I rather like and is where this package takes its name.

The Wooldridge ETWFE solution is intuitive and elegant. But it is also rather
tedious and error prone to code up manually. The **etwfe** package aims to
simplify the process by providing convenience functions that do the work for
you.

## Dataset

To demonstrate the core functionality of **etwfe**, we’ll use the
[`mpdta`](https://bcallaway11.github.io/did/reference/mpdta.html) dataset on 
US teen employment from the **did** package (which you’ll need to install
separately).

```{r}
# install.packages("did")
data("mpdta", package = "did")
head(mpdta)
```

In the examples that follow, our goal is to estimate the effect of treatment
(`treat`) on log employment (`lemp`). Notice that the panel ID is at the county
level (`countyreal`), but treatment was staggered across cohorts (`first.treat`)
so that a group of counties were treated at the same time. In addition to these
staggered treatment effects, we also observe log population (`lpop`) as a 
potential, time-invariant control.^[The time-invariance property is important;
time-varying controls are not allowed in the ETWFE framework. But you can also
omit controls if you don't have any.]

## Basic usage

Okay, let's load **etwfe** and work through its basic functionality.
Given the package name, it won't surprise you to know that the key estimating
function is `etwfe()`. Here's how it would look for our example dataset.

```{r}
library(etwfe)

mod =
  etwfe(
    fml  = lemp ~ lpop, # outcome ~ (time-invariant) controls
    tvar = year,        # time variable
    gvar = first.treat, # group variable
    data = mpdta,       # dataset
    vcov = ~countyreal  # vcov adjustment (here: clustered)
    )
```

There are a few things to say about our `etwfe()` argument choices and other
function options, but we'll leave those details aside until a bit later. Right
now, just know that all of the above arguments are required except `vcov`
(though I generally recommend it too, since we probably want to cluster our
standard errors at the individual unit level). 

Let's take a look at our model object.
```{r}
mod
```

What `etwfe()` has done underneath the hood is construct a treatment dummy
variable `.Dtreat` and saturated it together with the other variables of
interest as a set of multiway interaction terms.^[It has also demeaned the 
`lpop` control variable, but that again is something we'll
get back too later. It's not particularly important for interpreting the final
results.] 

You may also have noticed that our `etwfe()` call returns a standard 
[**fixest**](https://lrberge.github.io/fixest/) object, since this is what it
uses to perform the underlying estimation. All of the associated methods and
functions from the
**fixest** package are thus compatible with our model object. For example, we
could plot the raw regression coefficients with 
`fixest::coefplot()`, or print them to a nice regression table with
`fixest::etable()`.  However, the raw coefficients from an `etwfe()` estimation
are not particularly meaningful in of themselves. Remember that these are complex
multiway interaction terms that are probably hard to to interpret on their own.
So, we probably want to aggregate them along some dimension of interest instead
(e.g., by groups or time, or as an event study). A natural way to perform these
aggregations is by recovering the appropriate marginal effects. The **etwfe**
package provides another convenience function for doing so: `emfx()`, which is a
thin(ish) wrapper around `marginaleffects::marginaleffects()`.

For example, we can recover the average treatment effect as follows.

```{r}
emfx(mod)
```

Or, we can recover the dynamic treatment effects _a la_ an event study by invoking
the type argument. This time I'll save the resulting object since I plan to reuse it
in a moment.

```{r}
mod_es = emfx(mod, type = "event")
# mod_es
summary(mod_es) # Summary print looks a bit nicer
```

(Other types of aggregations include `"calendar"` (here: treatment by year effects) 
and `"group"` (here: treatment by cohort).)

Since `emfx` produces a standard `marginaleffects` object, we can pass it on to
other supported methods and packages. For example, we can pass it on to
[**modelsummary**](https://vincentarelbundock.github.io/modelsummary/index.html)
to get a nice regression table of the event study coefficients. Note the use of
the `shape` and `coef_rename` arguments below; these are optional but help to
make the output look a bit nicer.

```{r}
library(modelsummary)

rename_fn = function(old_names) {
  new_names = gsub(".Dtreat", "Years post treatment =", old_names)
  setNames(new_names, old_names)
}

modelsummary(
  mod_es,
  shape = term:event:statistic ~ model,
  coef_rename =  rename_fn,
  gof_omit = "Adj|Within|IC|RMSE",
  title = "Event study",
  notes = "Std. errors are clustered at the county level"
)
```
For visualization, you can pass it on to your preferred plotting method. For example:

```{r}
library(ggplot2)

ggplot(mod_es, aes(x = event, y = dydx, ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0, col = "grey50") +
  geom_pointrange() +
  labs(x = "Years post treatment", y = "Effect on log employment") +
  theme_minimal()
```

## Other families

A key feature of the ETWFE approach---one that sets it apart from other advanced
DiD implementations and extensions---is that it supports nonlinear model
(distribution / link) families. Users need simply invoke the `family` argument.
Here's a brief example, where we recast our earlier event-study as a Poisson
regression.

```{r, warning=FALSE, message=FALSE}
mpdta$emp = exp(mpdta$lemp)

etwfe(
  emp ~ lpop, tvar = year, gvar = first.treat, data = mpdta, vcov = ~countyreal,
  family = "poisson"
  ) |>
  emfx("event") |>
  summary()
```

## What is etwfe doing under the hood?

Now that you've seen **etwfe** in action, let's circle back to what the package
is doing under the hood. This section isn't necessary for you to use any of the 
functions; feel free to skip it. But a review of the internal details should help
you to optimize for different scenarios and also give you an understanding of
**etwfe's** default choices. 

As I've stated several times, the Wooldridge ETWFE approach basically involves
saturating the regression with interaction effects. You can easily grab the
formula of an estimated model to see for yourself.

```{r}
mod$fml_all
```

At this point, however, you may notice a few things. The first is that our
formula references several variables that aren't in the original dataset. An
obvious one is the `.Dtreat` treatment dummy. A more subtle one is `lpop_dm`,
which is the _demeaned_ (i.e., group-centered) version of our `lpop` control
variable.  All (time-invariant) control variables have to be demeaned before
they are interacted in the ETWFE setting. Here's how you could have constructed
the dataset ahead of time and estimated the ETWFE regression manually:

```{r}
# First construct the dataset
mpdta2 = mpdta |>
  transform(
    .Dtreat = year >= first.treat & first.treat != 0,
    lpop_dm = ave(lpop, first.treat, year, FUN = \(x) x - mean(x, na.rm = TRUE))
  )

# Then estimate the manual version of etwfe
mod2 = fixest::feols(
  lemp ~ .Dtreat:i(first.treat, i.year, ref = 0, ref2 = 2003) / lpop_dm |
    first.treat[lpop] + year[lpop],
  data = mpdta2,
  vcov = ~countyreal
)
```

We can confirm that the manual approach yields the same output as our original
etwfe regression. I'll use `modelsummary` to do that here, since I've already
loaded it above.^[Another option would be to use `fixest::etable()`].

```{r}
modelsummary(
  list("etwfe" = mod, "manual" = mod2),
  gof_map = NA # drop all goodness-of-fit info for brevity
)
```

Let's switch gears and talk about fixed-effects quickly. If you are a regular 
**fixest** user, you may have noticed that we've been invoking its 
[varying
slopes](https://lrberge.github.io/fixest/articles/fixest_walkthrough.html#varying-slopes-fex)
syntax in the fixed-effect slot (i.e., `first.treat[lpop]` and `year[lpop]`).
The reason for this is part practical, part philosophical. From a practical
perspective, `factor_var[numeric_var]` is equivalent to base R's 
`factor_var/numeric_var` "nesting" syntax but is much faster for high-dimensional 
factors.^[We won't see a speed-up for this small dataset, but it can make a
significant difference for large datasets.] From a philosphical perspective,
**etwfe** tries to limit the amount of extraneous information that it reports
to users. Most of the interaction effects in the ETWFE framework are just acting
as controls. By relegating them to the fixed-effects slot, we can avoid polluting
the user's console with a host of extra coefficients. Nonetheless, we can control 
this behaviour with the `fe` ("fixed effects") argument. Consider the following
options and their manual equivalents.

```{r, message=FALSE, warning=FALSE}
# fe = "feo" (fixed effects only)
mod_feo = etwfe(
  lemp ~ lpop, tvar = year, gvar = first.treat, data = mpdta, vcov = ~countyreal,
  fe = "feo"
)
# ... which is equivalent to the manual regression
mod_feo2 = fixest::feols(
  lemp ~ .Dtreat:i(first.treat, i.year, ref = 0, ref2 = 2003) / lpop_dm +
    lpop + i(first.treat, lpop, ref = 0) + i(year, lpop, ref = 2003) |
    first.treat + year,
  data = mpdta2, vcov = ~countyreal
)

# fe = "none"
mod_none = etwfe(
  lemp ~ lpop, tvar = year, gvar = first.treat, data = mpdta, vcov = ~countyreal,
  fe = "none"
)
# ... which is equivalent to the manual regression
mod_none2 = fixest::feols(
  lemp ~ .Dtreat:i(first.treat, i.year, ref = 0, ref2 = 2003) / lpop_dm +
    lpop + i(first.treat, lpop, ref = 0) + i(year, lpop, ref = 2003) +
    i(first.treat, ref = 0) + i(year, ref = 2003),
  data = mpdta2, vcov = ~countyreal
)
```

I'll leave it up to you to pass any of these models to `emfx` to confirm that
they give correct aggregated treatment effects. But we can quickly demonstrate
in a regression table that they all return the same raw coefficients.

```{r}
mods = list(
  "etwfe"         = mod,
  "manual"        = mod2,
  "etwfe (feo)"   = mod_feo,
  "manual (feo)"  = mod_feo2,
  "etwfe (none)"  = mod_none,
  "manual (none)" = mod_none2
)

modelsummary(mods, gof_map = NA)
```
